---
title: "NZ Soils"
author: "Rachel Kaminsky"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, fig.width=15, fig.height=11, warning=FALSE, message=FALSE)
```

##Summary
This document contains R code for  for chapters 4 and 5 of my PhD thesis

###Load packages
```{r}
library(phyloseq)  
library(ggplot2)
library(ggthemes)
```

###Set working directory and add files
```{r}
setwd("/Your_FilePath")
otu <- paste("otu_table.biom", sep = "")
map <- paste("map.txt", sep = "")
otu <- import_biom(otu)
map <- import_qiime_sample_data(map)
nzs <- merge_phyloseq(map, otu)
nzs
```

###Remove OTUs that aren't present in at least one sample, confirm all taxa > 0
```{r}
nzs_rm = prune_taxa(taxa_sums(nzs) > 0, nzs)
any(taxa_sums(nzs_rm)== 0)
```

###Rename ranks
```{r}
colnames(tax_table(nzs_rm)) = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
```

###Divide counts by 10 and round
```{r}
nzs_10  = transform_sample_counts(nzs_rm, function(x) x / 10) 
nzsr  = transform_sample_counts(nzs_10, round )
```

###Statistics for chemical data
```{r}
##Make data frame from mapping file
nzs_map = data.frame(sample_data(nzsr))

##ANOVA
fit <- aov(Sulfate ~ Time_Point + Treatment, data=nzs_map)

##Post-hoc test to determine factors that drive significant changes
TukeyHSD(fit)

```

###Calculate and plot observed species and shannon by pH and land use
```{r}
p = plot_richness(nzsr, x = "LandUse", color = "pH", measures = c("Observed", "Shannon"), sortby = "Observed") + theme_few() + theme(text = element_text(size=25), axis.text.x = element_text(angle = 90, hjust = 1)) + scale_color_gradient(low = "thistle", high = "skyblue3") + geom_point(size = 5) + geom_boxplot()
p
```

###Calculate and plot observed species and shannon by time point, treatment and land use
```{r}
##Order time points
name_order <- c("May-14", "Nov-14", "May-15")
sample_data(nzsr)$TimePoint <- factor(sample_data(nzsr)$TimePoint, levels=name_order)

##Plot diversity metrics
p = plot_richness(nzsr, x = "LandUse", color = "Treatment", measures = c("Observed", "Shannon"), sortby = "Observed") + theme_minimal() + theme(text = element_text(size=25), axis.text.x = element_text(angle = 90, hjust = 1)) + scale_color_gradient(low = "thistle", high = "skyblue3") + geom_point(size = 5) + geom_boxplot() + facet_grid(~Time_Point)
p
```

###Calculate and add observed species and shannon to metadata
```{r}
rich = estimate_richness(nzsr, measures = c("Observed", "Shannon"))
rich = data.frame(rich, SampleID = sample_names(nzsr))
rich_map = data.frame(nzsr_map)
rich_meta = merge(rich, rich_map, by = "SampleID")

```

###Regress by pH
```{r}
Shannon = rich_meta$Shannon
pH = rich_meta$pH
plot(pH, Shannon, xlab = "pH", ylab = "Shannon", cex.lab = 1.5, cex.axis = 1.5, pch = 16)
lmshan <- lm(Shannon ~ pH)
abline(lmshan)
summary(lmshan)

Observed = rich_meta$Observed
plot(pH, Observed, xlab = "pH", ylab = "Observed")
lmobs <- lm(Observed ~ pH)
abline(lmobs)
summary(lmobs)
```

###Kruskal-Wallis test for correlation between observed species and shannon and land use 
```{r}
kruskal.test(Observed ~ LandUse, data = rich)
kruskal.test(Shannon ~ LandUse, data = rich)

##repeat for other categorical variables
```

###Create bray-curtis dm and plot ordination by pH and land use
```{r}
nzs_dca = ordinate(nzsr, "DCA", "bray")
p1 = plot_ordination(nzsr, tpa_dca, color = "pH", shape = "LandUse") + theme_few() + theme(text = element_text(size=25), axis.text.x = element_text(angle = 90, hjust = 1)) + scale_color_gradient(low = "thistle", high = "skyblue3") + geom_point(size = 5)
p1

##add facet_grid(~Time_Point) to add temporal component, change color to Treatment to analyze treatment effects.
```

###Hierarchical cluster analysis
```{r}
https://raw.githubusercontent.com/nielshanson/mp_tutorial/master/taxonomic_analysis/code/pvclust_bcdist.R

source("pvclust_bcdist.R")
nzsr_mat <- as.matrix(otu_table(nzsr))
library(pvclust)
fit <- pvclust(nzsr_mat, method.hclust="ward.D",
   method.dist="bray–curtis")
plot(fit)
```

###Supportive statistics for pH (Mantel) and caterogical variables (Anosim)
```{r}
#Mantel
nzsr_dist <- distance(nzsr, "bray")
nzsr_sd <- data.frame(sample_data(nzsr))
nzsr_ph <- dist(nzsr_sd$pH, method = "euclidean")
nzsr_phd <- as.matrix(nzsr_ph)
mantel(nzsr_dist, nzsr_phd, method = "pearson")

#Anosim
nzsr_group = get_variable(nzsr, "LandUse")
nzsr_ano = anosim(distance(nzsr, "bray"), nzsr_group)
nzsr_ano$signif
nzsr_ano$statistic
plot(nzsr_ano)

#repeat Anosim for other categorical variables
```

###Calculate and plot observed species and shannon by soil order and subgroup
```{r}
p = plot_richness(nzsr, x = "SoilOrder", measures = c("Observed", "Shannon"), sortby = "Observed") + theme_few() + theme(text = element_text(size=23), axis.text.x = element_text(angle = 90, hjust = 1)) + geom_point(size = 5) + geom_boxplot()
z= p+ theme(axis.title.x = element_blank(),
             axis.title.y = element_text(size=27), strip.text.x = element_text(size=32))

p = plot_richness(nzsr, x = "SoilSubgroup", measures = c("Observed", "Shannon"), sortby = "Observed") + theme_few() + theme(text = element_text(size=23), axis.text.x = element_text(angle = 90, hjust = 1)) + geom_point(size = 5) + geom_boxplot()
z= p+ theme(axis.title.x = element_blank(),
             axis.title.y = element_text(size=27), strip.text.x = element_text(size=32))
```

###Calculate and add observed species and shannon to metadata
```{r}
rich = estimate_richness(nzsr, measures = c("Observed", "Shannon"))
rich = data.frame(rich, SampleID = sample_names(tpar))
rich_map = data.frame(tpa_map)
rich_meta = merge(rich, rich_map, by = "SampleID")
```

###Kruskal-Wallis test for correlation between observed species and shannon and soil order and subgroup 
```{r}
kruskal.test(Observed ~ SoilOrder, data = rich)
kruskal.test(Shannon ~ SoilOrder, data = rich)

kruskal.test(Observed ~ SoilSubgroup, data = rich)
kruskal.test(Shannon ~ SoilSubgroup, data = rich)
```

###Subset data by land use
```{r}
hc=as.vector(sample_data(nzsr) $LandUse=="High Country")
hcp=prune_samples(hc,nzsr)
hcp=prune_taxa(taxa_sums(hcp)>0, hcp)

sb=as.vector(sample_data(nzsr) $LandUse=="Sheep+Beef")
sbp=prune_samples(sb,nzsr)
sbp=prune_taxa(taxa_sums(sbp)>0, sbp)

dairy=as.vector(sample_data(nzsr) $LandUse=="Dairy")
dairyp=prune_samples(dairy, nzsr)
dairyp=prune_taxa(taxa_sums(dairyp)>0, dairyp)
```

###Subset data by region
```{r}
nc=as.vector(sample_data(nzsr) $Region=="North Canterbury")
ncp=prune_samples(nc, nzsr)
ncp=prune_taxa(taxa_sums(ncp)>0, ncp)

sc=as.vector(sample_data(nzsr) $Region=="South Canterbury")
scp=prune_samples(sc, nzsr)
scp=prune_taxa(taxa_sums(scp)>0, scp)

otago=as.vector(sample_data(nzsr) $Region=="Otago")
otagop=prune_samples(otago, nzsr)
otagop=prune_taxa(taxa_sums(otagop)>0, otagop)

south=as.vector(sample_data(nzsr) $Region=="Southland")
southp=prune_samples(south, nzsr)
southp=prune_taxa(taxa_sums(southp)>0, southp)
```

###Create ordinations for land use and region subgroups
```{r}
sb_dca = ordinate(sbp, "DCA", "bray")
hc_dca = ordinate(hcp, "DCA", "bray")
dairy_dca = ordinate(dairyp, "DCA", "bray")
nc_dca = ordinate(ncp, "DCA", "bray")
sc_dca = ordinate(scp, "DCA", "bray")
south_dca = ordinate(southp, "DCA", "bray")
otago_dca = ordinate(otagop, "DCA", "bray")
```

###Plot ordinations
```{r}
p1 = plot_ordination(sbp, sb_dca, color = "SoilSubgroup", shape = "SoilOrder") + theme_few() + theme(text = element_text(size=23), axis.text.x = element_text(angle = 90, hjust = 1)) + scale_color_manual(values = c("orangered", "indianred2", "mediumpurple", "violetred2")) + geom_point(size = 5) + ggtitle("Sheep+Beef")
p2 = plot_ordination(hcp, hc_dca, color = "SoilSubgroup", shape = "SoilOrder") + theme_few() + theme(text = element_text(size=23), axis.text.x = element_text(angle = 90, hjust = 1)) + scale_color_manual(values = c("red4", "slateblue", "sienna1", "mediumorchid1", "rosybrown2")) + geom_point(size = 5) + ggtitle("High Country")
p3 = plot_ordination(dairyp, dairy_dca, color = "SoilSubgroup", shape = "SoilOrder") + theme_few() + theme(text = element_text(size=23), axis.text.x = element_text(angle = 90, hjust = 1)) + scale_color_manual(values = c("slateblue4", "steelblue1", "seagreen3", "lightpink", "turquoise4", "mediumorchid1", "turquoise1",   "seagreen")) + geom_point(size = 5) + ggtitle("Dairy")
p4 = plot_ordination(ncp, nc_dca, color = "SoilSubgroup", shape = "SoilOrder") + theme_few() + theme(text = element_text(size=23), axis.text.x = element_text(angle = 90, hjust = 1)) + scale_color_manual(values = c("slateblue4", "steelblue1", "sienna1", "rosybrown2", "turquoise1")) + geom_point(size = 5) + ggtitle("North Canterbury")
p5 = plot_ordination(scp, sc_dca, color = "SoilSubgroup", shape = "SoilOrder") + theme_few() + theme(text = element_text(size=23), axis.text.x = element_text(angle = 90, hjust = 1)) + scale_color_manual(values = c("red4", "orangered", "indianred2", "lightpink", "turquoise4", "sienna1")) + geom_point(size = 5) + ggtitle("South Canterbury")
p6 = plot_ordination(southp, south_dca, color = "SoilSubgroup", shape = "SoilOrder") + theme_few() + theme(text = element_text(size=23), axis.text.x = element_text(angle = 90, hjust = 1)) + scale_color_manual(values = c("mediumpurple", "slateblue", "seagreen3", "violetred2", "mediumorchid1", "rosybrown2")) + geom_point(size = 5) + ggtitle("Southland")
p7 = plot_ordination(otagop, otago_dca, color = "SoilSubgroup", shape = "SoilOrder") + theme_few() + theme(text = element_text(size=23), axis.text.x = element_text(angle = 90, hjust = 1)) + scale_color_manual(values = c("mediumpurple", "slateblue", "seagreen3", "mediumorchid1", "seagreen")) + geom_point(size = 5) + ggtitle("Otago")

## add facet_grid(~Time_Point) to add temporal component, change color to Treatment to analyze treatment effects.
```

###Supportive statistics for soil order and soil subgroup (Anosim)
```{r}
#SB
sbp_group = get_variable(sbp, "SoilOrder")
sbp_ano = anosim(distance(sbp, "bray"), sbp_group)
sbp_ano$signif
sbp_ano$statistic

sbp_group = get_variable(sbp, "SoilSubgroup")
sbp_ano = anosim(distance(sbp, "bray"), sbp_group)
sbp_ano$signif
sbp_ano$statistic
plot(sbp_ano)

#repeat for other land use and region subgroups, and for treatment and time point
```


###Find taxa that change significantly between high and low country soils
```{r}
library("DESeq2")
sample_data(nzsr)$High.vs.Low <- relevel(sample_data(tpar)$High.vs.Low, "Low Country")
diagdds = phyloseq_to_deseq2(tpar, ~High.vs.Low)
diagdds = DESeq(diagdds,test="Wald", fitType = "local")
res = results(diagdds, cooksCutoff = FALSE)
res = res[order(res$padj, na.last = NA), ]
alpha = 0.001
sigtab = res[(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(tpar)[rownames(sigtab), ], "matrix"))
posigtab = sigtab[sigtab[, "log2FoldChange"] > 0, ]
posigtab = sigtab[sigtab[, "stat"] > 6, ]
posigtab = posigtab[, c("baseMean", "log2FoldChange", "lfcSE", "stat", "padj", "Phylum", "Class", "Family", "Genus","Species")]
```

###Plot high versus low significant taxa
```{r}
theme_set(theme_few())
scale_fill_discrete <- function(palname = "Set1", ...) {
  scale_fill_brewer(palette = palname, ...)
}
sigtabgen = subset(posigtab, !is.na(Genus))
x = tapply(sigtabgen$stat, sigtabgen$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtabgen$Phylum = factor(as.character(sigtabgen$Phylum), levels = names(x))
x = tapply(sigtabgen$stat, sigtabgen$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtabgen$Genus = factor(as.character(sigtabgen$Genus), levels = names(x))
write.csv(sigtabgen, "Sigtabgen.csv")
sigtabgen2 = read.csv("Sigtabgen.csv", head=T)
ggplot(sigtabgen2, aes(x = reorder(Genus,-stat,sum), y = stat, color = Phylum)) + geom_point(size = 6) + theme(text = element_text(size=25)) +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = 0.5))
```

###Make dataframe containing significant taxa with metadata for pH
```{r}
taxonomy = tax_table(tpar, errorIfNULL=FALSE)
if( !is.null(taxonomy) ){
  taxonomy = data.frame(as(taxonomy, "matrix"))
} 

otus <- otu_table(tpar)
otus_mat <- otus@.Data
variable <- sample_data(tpar)
variable <- as.data.frame(variable, rownames.force = NA)
variable <- as.data.frame(apply(variable, 2, as.numeric))
results <- apply(otus_mat, 1 , function(x)  cor.test(variable$pH,x, method = "spearman")[c(3,4)])
results_df <- as.data.frame(results, rownames.force = NA)
p.vals <- unname((unlist(results)))[c(seq(from = 1, to = 230890, by = 2))]
p.vals_df <- as.data.frame(p.vals, rownames.force = NA)
phos <- unname((unlist(results)))[c(seq(from = 2, to = 230890, by = 2))]
phos_df <- as.data.frame(phos, rownames.force = NA)
fdr_adj <- p.adjust(p.vals,method = "BH")
fdr_adj_df <- as.data.frame(fdr_adj, rownames.force = NA)
Spearmans = cbind(as(tax_table(taxonomy), "matrix"), as(fdr_adj_df, "data.frame"))
Spearmans = cbind(as(Spearmans, "data.frame"), as(phos_df, "data.frame"))
Spearmans = cbind(as(Spearmans, "data.frame"), as(p.vals_df, "data.frame"))
colnames(Spearmans) = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species", "pval_adj_BH", "Rho", "p.value")
library(dplyr)
Spearmans_sort <- arrange(Spearmans, pval_adj_BH)
Spearman_mostsig <- subset(Spearmans, Rho >= 0.6 & pval_adj_BH < 0.001 | Rho <= -0.6 & pval_adj_BH < 0.001)
```

###Plot significant taxa for pH
```{r}
pHsigtabgen = subset(Spearman_mostsig, !is.na(Genus))
x = tapply(pHsigtabgen$Rho, pHsigtabgen$Phylum, function(x) max(x))
x = sort(x, TRUE)
pHsigtabgen$Phylum = factor(as.character(pHsigtabgen$Phylum), levels = names(x))
x = tapply(pHsigtabgen$Rho, pHsigtabgen$Genus, function(x) max(x))
x = sort(x, TRUE)
pHsigtabgen$Genus = factor(as.character(pHsigtabgen$Genus), levels = names(x))
write.csv(pHsigtabgen, "pHSigtabgen.csv")
pHsigtabgen2 = read.csv("pHSigtabgen.csv", head=T)
ggplot(pHsigtabgen2, aes(x = reorder(Genus,-Rho,sum), y = Rho, color = Phylum)) + geom_point(size = 6) + theme(text = element_text(size=25)) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = 0.5))
```

###Can also use ALDEx2 for finding significant taxa for all variables
```{r}
nzsr_dat <- data.frame(otu_table(nzsr))
nzsr_sam <- data.frame(sample_data(nzsr))

library(ALDEx2)
nzsr.aldex.clr <- aldex.clr(nzsr, mc.samples = 128)
nzsr.lu.aldex.glm <- aldex.glm(nzsr.aldex.clr, nzsr_sam$LandUse)
nzsr.tp.aldex.glm <- aldex.glm(nzsr.aldex.clr, nzsr_sam$Time_Point)
nzsr.tx.aldex.glm <- aldex.glm(nzsr.aldex.clr, nzsr_sam$Treatment)
nzsr.so.aldex.glm <- aldex.glm(nzsr.aldex.clr, nzsr_sam$SoilOrder)
nzsr.pH.aldex.corr <- aldex.corr(nzsr.aldex.clr, nzsr_sam$pH)
```